---
title: "Cleaning and exploring rice production, import and export data of Bangladesh"
author: "Wahiduzzaman Khan"
output:
  html_document:
    df_print: paged
---
In this project we are going to take a look at the production, import and export of rice of Bangladesh.</br>
You can download these data sets from [here](https://google.com). These data are collected from BRRI (Bangladesh Rice Research Institute) and WRS (World Rice Statistics).

First we are going load all the packages that we are going to use:

```{r}
library(tidyverse)
library(stringr)
library(readxl)
```

As all the packages are loaded, its time to read in the data. The first data file that we are going to read is the "rice production.xlsx". Opening this excel file we can see that there are 10 columns containg information about the harvested area, production and yield of the three variants of rice (Aus, Aman, Boro) for years 1971 to 2017. Lets import this file and store it as "RiceProduction".

```{r}
RiceProduction <- read_excel("raw data/rice production.xlsx", skip = 2, col_names = F)
```

While Imporing the data I skipped the first two rows because of their formatting. The way those two rows are formatted they are going to make a mess if they are imported. I also specified that the imported rows do not conatin column names. So now lets take a look at the imported data.

```{r}
glimpse(RiceProduction)
```

We can see that the is imported correctly but the names of the variables are randomly given and this kind of names make the data hard to work with. So lets give them appropiate names that explains what information they contain.

```{r}
colnames(RiceProduction) <- c("Year", "AreaAus", "AreaAman", "AreaBoro", "ProAus", "ProAman", "ProBoro", "YieldAus", "YieldAman", "YieldBoro")
glimpse(RiceProduction)
```

This alredy looks much better. In this dataset, the areas cultivated are expressed in 1000 hactors and produced amounts are expressed in 1000 tons. But for analysis we want them to show actuall amount. So we have to multiply those columns by 1000. To do this we are going to create a function called "multiplier".

```{r}
multiplier <- function(table, i, n){
    table_final <- table
    for (i in i) {
        table_final[,i] <- table_final[,i] * n
    }
    return(table_final)
}
```

This function takes a dataset, column numbers and number as arguments and multiplies all those columns by the number.

```{r}
RiceProduction <- multiplier(RiceProduction, c(2,3,4,5,6,7), 1000)
```

Now we have the actual values.</br>
Lets put our concentrarion on the Year column now. We can see it has a range of years instead of a single year. In this analysis we are going to consider the lower bound as the year and remove the upper bound. We are also going to change Year column from character type to integers. We accomplish this by running the following code:

```{r}
RiceProduction <- separate(RiceProduction, Year, c("Year"), sep = "-")
RiceProduction$Year <- as.integer(RiceProduction$Year)
glimpse(RiceProduction)
```

We can see the Year column has a single year in each row. 